---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  # collapse = TRUE,
  # comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CRediTas

<!-- badges: start -->
[![R-CMD-check](https://github.com/ropensci/CRediTas/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ropensci/CRediTas/actions/workflows/R-CMD-check.yaml)
[![Status at rOpenSci Software Peer Review](https://badges.ropensci.org/576_status.svg)](https://github.com/ropensci/software-review/issues/576)
<!-- badges: end -->

The goal of CRediTas is to facilitate the tedious job of creating [CRediT authors statements](https://credit.niso.org/) for scientific publications. Normally, the first author of a scientific paper organizes a table in a spreadsheet where all the authors self-state their contributions. Often too, it is the first author responsibility to state the contributions of all co-authors. However, at the end, the information has to be translated to the CRediT statement format of "Author 1: roles Authors 2: roles ..." which is prone to errors and tedious, especially if there are many co-authors. The CRediTas package aims to make this easier by providing a template to be filled in form of a table (csv) and then converting this table to CRediT statement format.  

## Installation

You can install the development version of CRediTas from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("ropensci/CRediTas")
```

## Example

The workflow is meant to work with three basic functions. First, we create a template table. It can be created as a `data.frame` and being populated in R. Or as a csv file and being populated in your preferred csv editor.

```{r create_template}
library(CRediTas)

template_create(authors = c("Alexander Humboldt", "Carl Ritter"), file = tempfile())

cras_table <- template_create(authors = c("Friedrich Ratzel", 
                                          "Pau Vidal de la Blache", 
                                          "Pau Vila",
                                          "Élisée Reclus"))
knitr::kable(cras_table)

```
As you can see, the table is empty. So you must provide the information of who did what. You can use the `fix` function to fill the template directly in R:

```{r fix, eval=FALSE}

fix(cras_table)

```

If you wrote the template to a file, then you can read it back to R as follows:

```{r template_read, eval=FALSE}

cras_table <- template_read(path_to_your_csv_file)

```
Once the `cras_table` is populated, for instance:

```{r populate_random, echo=FALSE}

cras_table[-3, -1] <- sample(0:1, size=3*14, replace = TRUE, prob = c(0.6, 0.4))
knitr::kable(cras_table)

```
A text file can be generated following the CRediT author statement format. Since `drop = TRUE` by default, the authors without contribution are removed from the statement, Pau Vila in this case.

```{r}

textfile <- tempfile()

cras_write(cras_table, textfile, markdown = TRUE, quiet = TRUE)

```

If you open the text file, you will find this:

`r cras_write(cras_table, markdown = TRUE, quiet = TRUE)`


## Related packages

- [tenzing](https://github.com/marton-balazs-kovacs/tenzing/): Tenzing, an easy-to-use web-based app, allows researchers to generate reports about the contribution of each team member on a project using CRediT, for insertion into their manuscripts and for publishers to potentially incorporate into article metadata.
- [contribution](https://github.com/openbiox/contribution): The goal of contribution is to generate contribution table for credit assignment in a project. This is inspired by Nick Steinmetz (see twitter https://twitter.com/SteinmetzNeuro/status/1147241138291527681).
