---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  # collapse = TRUE,
  # comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# credit

<!-- badges: start -->
[![R-CMD-check](https://github.com/jospueyo/credit/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jospueyo/credit/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of credit is to facilitate the tedious job of creating [CRediT authors statements](https://credit.niso.org/) for scientific publications.

## Installation

You can install the development version of credit from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jospueyo/credit")
```

## Example

The workflow is meant to work with three basic functions. First, we create a template table. It can be created as a `data.frame` and being populated in R. Or as a csv file and being populated in your preferred csv editor.

```{r create_template}
library(credit)

create_template(authors = c("Alexander Humboldt", "Carl Ritter"), file = tempfile())

cras_table <- create_template(authors = c("Friedrich Ratzel", "Pau Vidal de la Blache", "Élisée Reclus"))
knitr::kable(cras_table)

```
As you can see, the table is empty. So you must provide the information of who did what. If you wrote the template to a file, then you can read it back to R as follows:

```{r read_template, eval=FALSE}

cras_table <- read_template(path_to_your_csv_file)

```
Once the `cras_table` is populated, for instance:

```{r populate_random, echo=FALSE}

cras_table[, 2:ncol(cras_table)] <- sample(0:1, size=3*14, replace = TRUE, prob = c(0.6, 0.4))
knitr::kable(cras_table)

```
A text file can be generated following the CRediT author statement format.

```{r}

textfile <- tempfile()

write_cras(cras_table, textfile, markdown = TRUE)

```

If you open the textfile, you will find this:

`r readLines(textfile)`

